Supplemental Information

### 2021-02-12 supp_v1.Rmd
message

---
title: "Supplemental Information"
author: "Craig Brinkerhoff"
date: "January 28, 2021"
output: word_document
---

@@

#Supplemental Information
Text S1
Tsivoglou & Neal (1976) outlined an energy dissipation model for the reaeration coefficient (K2, or k without depth correction) that is frequently generalized and used to scale k600 (Raymond et al. 2012; Ulseth et al. 2019). They showed that K2 scales linearly with the energy dissipation rate eD. Their derivation explicitly shows that eD=dH/dT for some segment of stream where H is water surface elevation and T is hydraulic residence time. For a stream segment with residence time T, the loss in water surface elevation is synonymous equivalent to the downward component of flow velocity V in the h dimension over the stream segment length (i.e. Sh). So, we can generalize this to dH/dT=V_h S_h≈VS_h. Finally, this can be converted to an energy dissipation rate per unit weight of water such that eD [L^2/T^3 ]=g dH/dT≈〖gVS〗_h. This model is convenient for our use because SWOT will only measure Sh and not channel slope S0. Further, Manning’s equation under non-uniform flow conditions requires Sh and not S¬0 and so we can conveniently merge the slope terms across equations and directly measure them from SWOT.

To my knowledge, most k600 upscaling studies to date have assumed uniform flow conditions (i.e. S_0=S_h) to train their upscaling parameters. This is presumably because channel slope is readily available in hydrographic data products while water surface slope is not. Therefore, future work should look at the influence of non-uniform flow on the upscaling of k600. Once SWOT launches, this will be easy to compare as SWOT data products will have both channel slope and water surface slope.

@@

![Figure S3 By-river performance metrics for FCO2 using the metrics outlined in Table 1. Dashed lines denote scores of 1, 0, and -0.41 for KGE (Knoben et al. 2019). Recall that the Raymond and Lauerwald models use the in-situ discharge record while BIKER does not.](outputs/flux_implications/FCO2_by_river.jpg)

@@

>>>>

---
title: "Supplemental Information"
author: "CB Brinkerhoff, CJ Gleason, PA Raymond, MH Harlan"
date: "`r Sys.Date()`"
csl: APA.csl
output:
  word_document: 
      reference_docx: word-styles-reference.docx
bibliography: References.bib

---

@@

##Text S1
[@tsivoglouTracerMeasurementReaeration1976] outlined an energy dissipation model for the reaeration coefficient ($K_2$, or k without depth correction) that is frequently generalized and used to scale $k_{600}$ [@ulsethDistinctAirWater2019a; @raymondScalingGasTransfer2012a]. They showed that $K_2$ scales linearly with the energy dissipation rate eD. Their derivation explicitly shows that $eD=\frac{dH}{dT}$ for some segment of stream where H is water surface elevation and T is hydraulic residence time. For a stream segment with residence time T, the loss in water surface elevation is synonymous with the downward component of flow velocity V in the h dimension over the reach length (i.e. $S_h$). So, we can generalize this to $\frac{dH}{dT}=V_hS_h≈VS_h$. Finally, this can be converted to an energy dissipation rate per unit weight of water such that $eD  [\frac{L^2}{T^3}]=g\frac{dH}{dT}≈gVS_h$. This model is convenient for our use because SWOT will only measure $S_h$ and not channel slope $S_0$. Further, Manning’s equation under non-uniform flow conditions requires $S_h$ and not $S_0$ and so we can conveniently merge the slope terms across equations and directly measure them from SWOT.

##Text S2
In Figure S4 we show that bed roughness only exerts an influence on $k_{600}$ when rivers are incredibly steep. We do this using the 'Keulegan effective roughness height' for the channel bed, which is detailed in equation S1 below (in the form form [@dingmanAnalyticalDerivationAtastation2007b]). This model was first outlined in the 1930s by Keulegan and outlined explicitly in [@fergusonHydraulicsHydraulicGeometry1986a] and [@dingmanAnalyticalDerivationAtastation2007b] in the context of hydraulic geometry (noting that it is a more generalized form of the Manning's and Chezy's velocity~depth relations commonly used in the literature). The model arises from depth integrating the Prandl-von Karman equation for velocity in a turbulent boundary layer ($\frac{u}{u_\star}=2.5ln(\frac{30y}{k_s})$) [@[@fergusonHydraulicsHydraulicGeometry1986a]].

$$V=2.5(gDS)^{1/2}ln(\frac{11D}{y_r})   (S1)$$
$y_r [L]$ is the effective bed roughness height we solved for using the data from [@ulsethDistinctAirWater2019a] where D is channel depth [L], S is channel slope, g is gravitational acceleration $[M/T^2]$, and V is average flow velocity $[M/T]$. Because it is an 'effective/equivalent height', this does not necessairly reflect actual roughness elements in the river channel but is a reasonable way to quantify relative differences in bed roughness across data. It is convenient to use here because it provides an actual roughness height and is not simply a coefficient like the conductance coefficients in Manning's and Chezy's velocity~depth relations.

In Figure S4, this effective roughness height was plotted against $k_{600}$ and the data was progressively filtered for steeper and steeper rivers.

@@

![Figure S3. Scaling $k_{600}$ using an effective bed roughness term (see Text S2). This scaling relationship only exists for rivers with very steep slopes, here presented as greater than 0.1](outputs/k600/roughness.jpg)

@@

##References
